14th International Satisfiability Modulo Theories Competition (SMT-COMP 2019): Rules and Procedures

### 2020-04-01 rules/rules20.tex
some updates, especially new rules

\author{%
\todo{check}
Haniel Barbosa\\
Universidade Federal de Minas Gerais\\
Brazil\\
{\small\href{mailto:hbarbosa@dcc.ufmg.br}{\texttt{hbarbosa@dcc.ufmg.br}}}\\
\and
Jochen Hoenicke\\
Universit\"at Freiburg\\
Germany
{\small\href{mailto:hoenicke@informatik.uni-freiburg.de}{\texttt{hoenicke@informatik.uni-freiburg.de}}}\\
\and
Antti Hyvarinen (chair)\\
Universita della Svizzera italiana \\
Switzerland \\
{\small\href{mailto:antti.hyvaerinen@usi.ch}{\texttt{antti.hyvaerinen@usi.ch}}} \\
}

@@

\begin{description}
\item[March~1] Deadline for new benchmark contributions.
\item[May~1] Final versions of competition tools (e.g., benchmark
  scrambler) are made available.  Benchmark libraries are frozen.
\item[May~4] Deadline for first versions of solvers (for all tracks),
  including information about which tracks and divisions are being
  entered, and magic numbers for benchmark scrambling.
\item[May~18] Deadline for final versions of solvers, including
  system descriptions.
\item[May~19] Opening value of NYSE Composite Index used to compute
  random seed for competition tools. \todo{Okay?}
\item[July~5/6] SMT Workshop; end of competition, presentation of
  results.
\end{description}

@@

The rest of this document, revised from the previous
version,\footnote{Earlier versions of this document include
  contributions from Clark Barrett, Roberto Bruttomesso, David Cok,
  Sylvain Conchon, David D{\'e}harbe, Morgan Deters, Alberto Griggio,
  Liana Hadarean,
  Matthias Heizmann, Aina Niemetz, Albert Oliveras, Giles Reger, Aaron Stump,
	and Tjark Weber.}  describes the rules and competition procedures for
SMT-COMP~2020.
The principal changes from the previous competition rules are the following:
\begin{itemize}
  \item {\bf Portfolio Solvers are Forbidden.} \todo{check}
    This year we explicitly forbid the submission of so-called
    portfolio solvers as competition entries.  We acknowledge that
    judging what constitutes a portfolio SMT solver is difficult.
    However, we are not the first in history to face a similar
    challenge, and plan to apply an equally historical approach where
    we write down a ``law'', act as its ``judges'', and have a few
    ``precedents''.

    Here we define a portfolio solver as ``a solver using a combination of
    two or more sub-solvers, developed by different groups of authors, on
    the same component or abstraction of the input problem.'' For example, a
    solver using one subsolver to solve the ground abstraction of a
    quantified problem is allowed, while a solver using two or more
    subsolvers from different groups of authors is not.
    \todo{Something about  hybrid solvers?}

    \todo{rationale}

  \item {\bf Mandatory System Descriptions.}
    SMT-COMP entrants are now required to provide a short (1–2 pages)
    description of the system.
    \rationale{The main incentive for this
    change is twofold.  First, we want to improve transparency when submitted
    solvers are wrapper or derived tools according to the rules of the
    competition.  Second, we want to encourage documenting technical
    improvements that lead to the current results.}

  \item {\bf Limit the Number of Solvers.} \todo{NEW}
    This year solver development teams may only submit more than one
    entry if they are substantially different.  A justification must
    be provided for the difference.  We strongly encourage the teams
    to keep the number of solvers per team per category at at most
    two.

    \rationale{The organizers wishes to keep the total run-time of the
      competition small, in case of unforeseen problems that require
      to rerun the whole competition.  The idea in allowing up to two
      submissions is to encourage the development of new, experimental
      techniques by allowing an “alternative solver” to be submitted
      while keeping the competition manageable.}

  \item {\bf Naming Convention for Derived Tools.}
    This year, we require that a derived tool (a tool based on and extending
    another SMT solver) should follow the \emph{naming convention} {[name of
    base solver]-[my solver name]}.
    \rationale{Give credit where credit is due.}

  \item {\bf Combine stdout and stderr.}
    This year, output to the standard error channel is not ignored, but
    treated the same as output to the standard output channel.

    \rationale{Due to limitations of starexec, the standard error channel
      and standard output channel are not separated.  The previous year the
      organizers had to work around the limitation by wrapping the submitted
      solvers in a script that ignored stderr.  By changing the rule we avoid
      that the organizers have to change the submitted solvers.}

%  \item {\bf Incremental Track.}
%    In previous years, benchmarks were not eligible for the \inctrack
%    if their first \akey{check-sat} command had unknown status.  Similarly, the
%    trace executor used to send commands from the benchmark to the solver's
%    standard input channel stopped execution at the first \akey{check-sat}
%    command with unknown status. This year, benchmarks whose first
%    \akey{check-sat} command has unknown status are eligible for the
%    \inctrack and trace execution is only stopped before the solver
%    completes a benchmark if the execution runs into the time limit.
%    \rationale{The previous limitations were imposed by the trace
%    executor, which now has been extended to support executing solvers on
%    benchmarks beyond \akey{check-sat} commands with unknown status.}

  \item {\bf New \mvaltrack for QF\_BV.}
    This year, we introduce a new experimental \mvaltrack for the QF\_BV logic.
    The benchmarks for this track will include all eligible \emph{non-incremental}
    QF\_BV benchmarks with known satisfiable status in SMT-LIB.
    Participating solvers are required to support the \akey{get-model} command.
    \rationale{In many SMT applications, model generation is an essential
    feature.  Previously, none of the SMT-COMP tracks required model
    generation. One of the challenges is that the model format is not
    consistent across different solvers. While imposing a standard over all
    logics is challenging, there are several logics (e.g., QF\_BV) where it is
    straightforward.  In the future we hope to expand this track to other
    logics as a way of pushing for model standardization.  Since the QF\_BV
    division was the one with the largest number of participants in previous
    years,  we hope for a high number of participants in this track.  }

  \item {\bf New \challtrack.}
    This year, we introduce a new \challtrack.  This track will contain
    new challenging SMT-LIB benchmarks (with an emphasis on industrial
    applications) that are either unsolved or unsolved within some reasonable
    time limit, as indicated by the benchmark submitters.  We will also include
    benchmarks nominated by the community as challenging and of interest. In
    this track, solvers will run with a significantly longer time limit.
    \rationale{ Different application domains require different time limits.
    For example software verification traditionally requires much lower time
    limits, compared to hardware verification.  To reward solvers optimized for
    different use cases, we propose this new track and a new score for
    benchmarks solved within a very low time limit (see item ``New Scores in
    the \maintrack''). See Section~\ref{sec:exec:industry-challenge} for the
    logics included in the \challtrack for this year. }

  \item {\bf Time Limit.} The time limit per solver/benchmark pair is
    anticipated to be at most 20 minutes in the \maintrack, \inctrack,
    \ucoretrack and \mvaltrack.  For the \challtrack, it is anticipated to be
    at most 720 (12 hours) minutes.
    \rationale{The time limit for the Main Track is
    reduced to 20 minutes (down from 40 minutes in earlier years) \todo{rationale}.
    For the \challtrack, 12 hours seem to be a reasonable compromise for the
    expected number of participants and benchmarks submitted to this track.}

  \item {\bf Benchmark Selection.} \todo{update}
    Since 2015, the competition evaluated all
    solvers on all eligible benchmarks in SMT-LIB.  This year, we will
    use an alternative benchmark selection scheme that selects a subset
    of the eligible benchmarks. The benchmark selection will be random,
    but will guarantee inclusion of newly submitted benchmarks.  The upper
    bound for the size of each division is not fixed, but
    depends on the size of the corresponding logic.  Prior to this
    selection, we will remove all benchmarks in a division that were
    solved by all solvers (including non-competing solvers) in this
    division in under one second in last year's competition.
    \rationale{Evaluating solvers on all eligible benchmarks in SMT-LIB
    makes results more predictable and seems to be more of an evaluation
    than a competition.  In the Main track last year,
    78\% of the 258,741 benchmarks were solved by all supported solvers
    within the time limit (71\% within 1 second). In 7 (out of 46)
    logics, over 99\% of benchmarks were solved by all solvers.
    Removing `easy' (or at least `unsurprising') benchmarks attempts to
    shift the focus towards challenging benchmarks, in particular since
    we now use an alternative benchmark selection scheme.  It can be
    argued that the size of a logic in SMT-LIB can be seen as an
    indicator of its relevance.  We thus do not use a fixed upper bound
    for the size of a division in order to reflect the suggested
    importance of a division.}

  \item {\bf Scoring Scheme.}
    This year, we will abandon the weighted scoring scheme that was introduced
    for the Main Track and \ucoretrack in 2016. We will fall back to the
    scoring scheme based on the number of solved instances that was last used
    in 2015.
    \rationale{ Since 2016, the competition has used a scoring scheme based on
    benchmark weights in order to de-emphasize large benchmark families.  The
    scoring scheme is fairly complicated and not necessarily intuitive.  It
    further makes comparing results in papers with results from the competition
    more difficult.  A recent analysis of competition data from 2015-2018 for a
    report on the competition of these years (currently under submission to
    JSAT) suggests that benchmark families do not have a significant impact on
    the (weighted) scores  When determining winners for each division using
    the scoring scheme of 2015, the winners for only 7 divisions (out of 139)
    over all three years would have changed.  While producing these results we
    further noticed that the description of what constitutes a benchmark family
    (as stated in the rules documents) had been incorrectly interpreted by the
    scoring scripts in 2016-2018. After fixing this misinterpretation, only one
    single division winner changes - in 2017 AUFNIRA should have been won by
    CVC4 and not Vampire.}

  \item {\bf New Scores in the \maintrack.}
    This year, additionally to the separate scores given for sequential and
    parallel performance, we will reward three new scores in the \maintrack.
    The \emph{24-second score} will reward solving performance within a time
    limit of 24 seconds (wall clock time), the \emph{sat score} will reward
    performance on satisfiable instances, and the \emph{unsat score} will
    reward performance on unsatisfiable instances.
    \rationale{ Different application domains of SMT typically impose a wide
    range of time limits (from hours to seconds).  Previous time limits used in
    the competition were the same for all benchmarks and thus agnostic to the
    application domain of a benchmark.  The new \challtrack and the new
    24-second score try to address use cases on both extreme ends of the
    spectrum.  Further, in many cases an SMT application produces either mainly
    satisfiable or mainly unsatisfiable queries to the SMT solver.  The new sat
    and unsat scores are intended to reward solvers that implement specialized
    techniques and optimizations for either case.}

  \item {\bf Do Not Run Non-Competitive Divisions.}
    This year, we will not run non-competitive divions.
    \rationale{Evaluating solvers in non-competitive divisions is more in the
    spirit of an evaluation than a competition.}

  \item {\bf Experimental Strings Division.}
    The competition will feature experimental divisions for benchmarks that
    use strings. Participating solvers must implement the semantics of the
    current SMT-LIB draft for the theory of unicode strings
    (\url{http://smtlib.cs.uiowa.edu/theories-UnicodeStrings.shtml}).
    \rationale{Corresponding theories and benchmarks are expected to be added
    to SMT-LIB in the near future.}

  \item {\bf Competition-Wide Recognitions.}
    This year, we will replace the previous notion of competi\-tion-wide scoring
    as introduced for the Main Track in 2014 (and improved over the years)
    with recognitions for all tracks that do not directly compare divisions.
    \rationale{Competi\-tion-wide scoring was introduced as a metric to
    compare solver performance across all divisions. It is, by definition,
    biased towards solvers that enter a large number of divisions.}

@@

\todo{Portfolio Tool}

@@

\header{Attendance.}\todo{fix this!}
%
Submitters of an SMT-COMP entrant are not required (but encouraged) to be
physically present at the competition or the SMT Workshop to participate or
win.

@@

\header{Aborts and Unparsable Output.}
%
In all tracks except the \inctrack, any \texttt{success} outputs will be
ignored.  Solvers that exit before the time limit without reporting a result
(e.g., due to exhausting memory or crashing) \emph{and} do not produce output
that includes \texttt{sat}, \texttt{unsat}, \texttt{unknown} or other track
specific output as specified in the individual track sections e.g. unsat cores
or models, will be considered to have aborted.

@@

\header{Time Limit.}
This track will use a wall-clock time limit of 20 minutes per solver/benchmark
pair.\todo{we said 20 minutes, right?}

@@

\header{Input and output.}
Participating solvers will be connected to a trace executor, which
will incrementally send commands to the standard input channel of the
solver and read responses from both the standard output channel of the
solver.  The commands will be taken from an SMT-LIB benchmark script
that satisfies the requirements for incremental track scripts given in
Section~\ref{sec:logistics}.
%
Solvers must respond to each command sent by the trace executor with
the answers defined in the SMT-LIB format specification, that is, with
an answer of \texttt{sat}, \texttt{unsat}, or \texttt{unknown} for
\akey{check-sat} commands, and with a \texttt{success} answer for
other commands.
Solver must not write anything to the standard error channel. \todo{okay?}

@@

\header{Time Limit.}
This track will use a wall-clock time limit of 20 minutes per solver/benchmark
pair.\todo{20 min}

@@

\header{Result.}
The result of a solver is considered \emph{erroneous} if (i) the
response to the \akey{check-sat} command is \texttt{sat}, (ii) the
returned unsatisfiable core is not well-formed (e.g., contains names of
formulas that have not been asserted before), or (iii) the returned
unsatisfiable core is not, in fact, unsatisfiable.
%
If the solver replies \texttt{unsat} to \akey{check-sat} but gives no
response to \akey{get-model}, this is considered as no reduction, i.e.,
as if the solver would have returned the entire benchmark as an unsat
core.

>>>>

\author{%
Haniel Barbosa\\
Universidade Federal de Minas Gerais\\
Brazil\\
{\small\href{mailto:hbarbosa@dcc.ufmg.br}{\texttt{hbarbosa@dcc.ufmg.br}}}\\
\and
Jochen Hoenicke\\
Universit\"at Freiburg\\
Germany\\
{\small\href{mailto:hoenicke@informatik.uni-freiburg.de}{\texttt{hoenicke@informatik.uni-freiburg.de}}}\\
\and
Antti Hyvarinen (chair)\\
Universita della Svizzera italiana \\
Switzerland \\
{\small\href{mailto:antti.hyvaerinen@usi.ch}{\texttt{antti.hyvaerinen@usi.ch}}} \\
}

@@

\begin{description}
\item[March~1] Deadline for new benchmark contributions.
\item[April~20] Final versions of competition tools (e.g., benchmark
  scrambler) are made available.  Benchmark libraries are frozen.
\item[May~4] Deadline for first versions of solvers (for all tracks),
  including information about which tracks and divisions are being
  entered, and magic numbers for benchmark scrambling.
\item[May~18] Deadline for final versions of solvers, including
  system descriptions.
\item[May~19] Opening value of NYSE Composite Index used to compute
  random seed for competition tools. \todo{Okay?}
\item[July~5/6] SMT Workshop; end of competition, presentation of
  results.
\end{description}

@@

The rest of this document, revised from the previous
version,\footnote{Earlier versions of this document include
  contributions from Clark Barrett, Roberto Bruttomesso, David Cok,
  Sylvain Conchon, David D{\'e}harbe, Morgan Deters, Alberto Griggio,
  Liana Hadarean,
  Matthias Heizmann, Aina Niemetz, Albert Oliveras, Giles Reger, Aaron Stump,
	and Tjark Weber.}  describes the rules and competition procedures for
SMT-COMP~2020.
The principal changes from the previous competition rules are the following:
\begin{itemize}
  \item {\bf Portfolio Solvers are Forbidden.}
    This year we explicitly forbid the submission of so-called
    portfolio solvers as competition entries.  We acknowledge that
    judging what constitutes a portfolio SMT solver is difficult.
    However, we are not the first in history to face a similar
    challenge, and plan to apply an equally historical approach where
    we write down a ``law'', act as its ``judges'', and have a few
    ``precedents''.
    Here we define a portfolio solver as ``a solver using a combination of
    two or more sub-solvers, developed by different groups of authors, on
    the same component or abstraction of the input problem.'' For example, a
    solver using one subsolver to solve the ground abstraction of a
    quantified problem is allowed, while a solver using two or more
    subsolvers from different groups of authors is not.
    \rationale{The goal of the competition is to evaluate and improve
      the state of the art of SMT solving. By just combining existing
      solvers and running them in parallel it is possible to win the
      competition without contributing anything new.}

  \item {\bf Limit the Number of Solvers.}
    This year solver development teams may only submit more than one
    entry if they are substantially different.  A justification must
    be provided for the difference.  We strongly encourage the teams
    to keep the number of solvers per team per category at at most
    two.
%
    \rationale{The organizers wishes to keep the total run-time of the
      competition small, in case of unforeseen problems that require
      to rerun the whole competition.  The idea in allowing up to two
      submissions is to encourage the development of new, experimental
      techniques by allowing an ``alternative solver'' to be submitted
      while keeping the competition manageable.}

  \item {\bf Combine stdout and stderr.}
    This year, output to the standard error channel is not ignored, but
    treated the same as output to the standard output channel.
    Solvers should not write anything to the standard error channel.
%
    \rationale{Due to limitations of starexec, the standard error channel
      and standard output channel are not separated.  In the previous year, the
      organizers had to work around the limitation by wrapping the submitted
      solvers in a script that ignored stderr.  By changing the rule we avoid
      that the organizers have to change the submitted solvers.}

%  \item {\bf The \mvaltrack is extended to other logics.}
%    This year, we introduce a new experimental \mvaltrack for the logics QF\_IDL, QF\_RDL, QF\_LIA, QF\_LRA, QF\_LIRA.
%    The benchmarks for this track will include all eligible \emph{non-incremental}
%    QF\_BV benchmarks with known satisfiable status in SMT-LIB.
%    Participating solvers are required to support the \akey{get-model} command.
%    \rationale{In many SMT applications, model generation is an essential
%    feature.  Previously, none of the SMT-COMP tracks required model
%    generation. One of the challenges is that the model format is not
%    consistent across different solvers. While imposing a standard over all
%    logics is challenging, there are several logics (e.g., QF\_BV) where it is
%    straightforward.  In the future we hope to expand this track to other
%    logics as a way of pushing for model standardization.  Since the QF\_BV
%    division was the one with the largest number of participants in previous
%    years,  we hope for a high number of participants in this track.  }

  \item \todo{\challtrack}

  \item {\bf Time Limit.} The time limit per solver/benchmark pair is
    anticipated to be at most 20 minutes in the \maintrack, \inctrack,
    \ucoretrack and \mvaltrack.
    \rationale{The time limit for the Main Track is
      reduced to 20 minutes (down from 40 minutes in the previous years),
      to make running the competition more manageable}

  \item {\bf Benchmark Selection.}
    This year we reduce the percentage of the benchmarks that are run in the competition.
    \rationale{The goal is to make running the competition more managable.}

  \item {\bf The Strings Division is not experimental.}
    The last years competition introduced a experimental divisions for benchmarks that
    use strings. This year, the String division is no longer experimental.

    \rationale{The final specification of the string theory has been officially released.}

@@

\header{Portfolio Solver.}
%
A \emph{portfolio solver} is a solver using a combination of two or
more sub-solvers, developed by different groups of authors, on the
same component or abstraction of the input problem.  For example, a
solver using one subsolver to solve the ground abstraction of a
quantified problem is allowed, while a solver using two or more
subsolvers from different groups of authors is not.
Porfolio solver are in general \textbf{not allowed}.
If you are unsure if your tool is a portfolio solver according to this
definition and you feel that it should be allowed contact the
organizers of the SMT-COMP for clarification.

@@

%\header{Attendance.}
%
%Submitters of an SMT-COMP entrant are not required (but encouraged) to be
%physically present at the competition or the SMT Workshop to participate or
%win.

@@

\header{Aborts and Unparsable Output.}
%
In all tracks except the \inctrack, any \texttt{success} outputs will be
ignored.  Solvers that exit before the time limit without reporting a result
(e.g., due to exhausting memory or crashing) \emph{and} do not produce output
that includes \texttt{sat}, \texttt{unsat}, \texttt{unknown} or other track
specific output as specified in the individual track sections e.g. unsat cores
or models, will be considered to have aborted.  Note that there is no
distinction between output and error channel and tools should not write any
message to the error channel that could be misinterpreted as a wrong result.

@@

\header{Time Limit.}
This track will use a wall-clock time limit of 20 minutes per solver/benchmark
pair.

@@

\header{Input and output.}
Participating solvers will be connected to a trace executor, which
will incrementally send commands to the standard input channel of the
solver and read responses from both the standard output channel of the
solver.  The commands will be taken from an SMT-LIB benchmark script
that satisfies the requirements for incremental track scripts given in
Section~\ref{sec:logistics}.
%
Solvers must respond to each command sent by the trace executor with
the answers defined in the SMT-LIB format specification, that is, with
an answer of \texttt{sat}, \texttt{unsat}, or \texttt{unknown} for
\akey{check-sat} commands, and with a \texttt{success} answer for
other commands.
Solvers must not write anything to the standard error channel.

@@

\header{Time Limit.}
This track will use a wall-clock time limit of 20 minutes per solver/benchmark
pair.

@@

\header{Result.}
The result of a solver is considered \emph{erroneous} if (i) the
response to the \akey{check-sat} command is \texttt{sat}, (ii) the
returned unsatisfiable core is not well-formed (e.g., contains names of
formulas that have not been asserted before), or (iii) the returned
unsatisfiable core is not, in fact, unsatisfiable.
%
If the solver replies \texttt{unsat} to \akey{check-sat} but gives no
response to \akey{get-unsat-core}, this is considered as no reduction, i.e.,
as if the solver would have returned the entire benchmark as an unsat
core.
